---
title: "ETM 540 Week 1 Homework"
author: "Jordan Hilton"
date: "January 8, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (pander, quietly = TRUE) 
library (magrittr, quietly = TRUE) 
library (dplyr, quietly = TRUE) 
library (ROI, quietly = TRUE) 
library (ROI.plugin.glpk, quietly = TRUE) 
library (ompr, quietly = TRUE)
library (ompr.roi, quietly = TRUE)
```

## Problem 1.1

I installed packages quietly in the setup for this RMD, including pander, magrittr, dplyr, ROI, and ompr.

Let's start with creating table b, and then transposing it, as from the book:

```{r}
b<-matrix(c(1,2,3,4,5,6,7,8), ncol=4, dimnames=c(list(c("Row1", "Row2")), list(c("Col1", "Col2","Col3","Col4"))))
d<-t(b) 
pander(d, caption="Transposition of Matrix b")
```

We can see the transposed names. Let's change the names for the matrix so that they make sense:

```{r }
dimnames(d)<-c(list(c("Row 1", "Row 2","Row 3","Row 4")), list(c("Column 1", "Column 2")))
pander(d, caption="Fixed Names")
```

## Problem 1.2

To do this let's adapt the code used above to create a matrix by changing the number of columns to seven and naming appropriately:

```{r}
demand<-matrix(c(10,15,20,25,30,35,40), ncol=7, dimnames=c("WIDGET!!!", list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
pander(demand, caption="Weekly Demand")
```

## Problem 1.3

Let's create each row with seq as above, and then rbind them since we're learning about rbind:

```{r}
demandweek1<-matrix(seq(10, by=5,length.out=7), ncol=7, dimnames=c("Week 1", list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
demandweek2<-matrix(seq(45, by=5,length.out=7), ncol=7, dimnames=c("Week 2", list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
demandweek3<-matrix(seq(80, by=5,length.out=7), ncol=7, dimnames=c("Week 3", list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
demandweek4<-matrix(seq(115, by=5,length.out=7), ncol=7, dimnames=c("Week 4", list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
demandmonth<-rbind(demandweek1, demandweek2,demandweek3,demandweek4)
pander(demandmonth, caption="Weekly Demand for a Month")
```


You can also do this in one line as follows:

```{r}
demand2<-matrix(seq(10, by=5,length.out=28), ncol=7, byrow=TRUE, dimnames=c(list(c("Week 1","Week 2","Week 3","Week 4")), list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
pander(demand2, caption="Another Way")
```

## Problem 1.4

Let's use our existing week 1 for Widgets and rbind to a similar row for gadgets:

```{r}
widgetsweek1<-matrix(seq(10, by=5,length.out=7), ncol=7, dimnames=c("WIDGETS!", list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
gadgetsweek1<-matrix(seq(20, by=3,length.out=7), ncol=7, dimnames=c("GADGETS!", list(c("Monday", "Tuesday","Wednesday","Thursday", "Friday","Saturday","Sunday"))))
twoproducts<-rbind(widgetsweek1, gadgetsweek1)
pander(twoproducts, caption="Weekly Demand for a Month")
```

## Problem 2.1

### a (Formulation)

Using the same format of formulation as in the book, let's first clearly define our choices:

* Chairs = # of Chairs to Make
* Desks = # of Desks to Make
* Tables = # of Tables to Make
* Frames = # of Frames to Make

Now let's formulate the LP:

$$
 \begin{split}
 \begin{aligned}
    \text{Maximize  }   & 20*Chairs+14*Desks+3*Frames+16*Tables \\
    \text{subject to } & 6*Chairs+2*Desks+Frames+4*Tables \leq 1440 \\
                       & 8*Chairs+6*Desks+Frames+8*Tables \leq 1440 \\
                       & 6*Chairs+4*Desks+Frames+25*Tables \leq 2000 \\
                      & 7*Chairs+10*Desks+2*Frames+12*Tables \leq 1000 \\
                       & 40*Chairs+25*Desks+5*Frames+16*Tables \leq 9600 \\
                       & Chairs, Desks, Frames, Tables \geq 0 \\   
  \end{aligned} 
  \end{split} 
$$

### b (Implementation)

Let's implement the model we've described above, as done in the book:

```{r base_case_4var}
result <- MIPModel() %>%
  add_variable(Chairs, type = "continuous", lb = 0) %>%
  add_variable(Desks, type = "continuous",lb = 0) %>%
  add_variable(Frames, type = "continuous", lb = 0) %>%
  add_variable(Tables, type = "continuous", lb = 0) %>%
  
  set_objective(20*Chairs + 14*Desks + 3*Frames + 16*Tables, "max") %>%
  
  add_constraint(6*Chairs+2*Desks+Frames+4*Tables <= 1440) %>% #fabrication
  add_constraint(8*Chairs+6*Desks+Frames+8*Tables <= 1440) %>% #assembly
  add_constraint(6*Chairs+4*Desks+Frames+25*Tables <= 2000) %>% #machining
  add_constraint(7*Chairs+10*Desks+2*Frames+12*Tables <= 1000) %>% #painting
  add_constraint(40*Chairs+25*Desks+5*Frames+16*Tables <= 9600) %>% #wood
 
  
  solve_model(with_ROI(solver = "glpk"))

```

### c (Solution)

Let's examine the outputs from solving our MIPModel, starting with the solver status:

```{r base_status_4var}
print(solver_status(result))
```

Hooray, we've obtained an optimal solution! How much stuff should we make?

```{r base_solution_4var}
get_solution(result, Chairs)
get_solution(result, Desks)
get_solution(result, Frames)
get_solution(result, Tables)
```

I hope you like chairs! Let's ask the important question- how much money did we make?

```{r base_objecfunc_4var}
print(objective_value(result))
```

### c (Interpretation and discussion)

So it looks like our model is just telling us that it's not worth it to produce anything except chairs, which must mean that they  use our bottleneck resource more efficiently than the other products.


### d (Changing one parameter)

I think the easiest way to obtain a different result is just to tell our model that another product is suddenly much more profitable (say, set the profit on frames to 30). It would be more interesting though to identify our bottleneck resource and increase the supply of that so that our model tells us to make a mix of products. Let's see what happens when we increase the supply of painting hours:

```{r morepaint}
result2 <- MIPModel() %>%
  add_variable(Chairs, type = "continuous", lb = 0) %>%
  add_variable(Desks, type = "continuous",lb = 0) %>%
  add_variable(Frames, type = "continuous", lb = 0) %>%
  add_variable(Tables, type = "continuous", lb = 0) %>%
  
  set_objective(20*Chairs + 14*Desks + 3*Frames + 16*Tables, "max") %>%
  
  add_constraint(6*Chairs+2*Desks+Frames+4*Tables <= 1440) %>% #fabrication
  add_constraint(8*Chairs+6*Desks+Frames+8*Tables <= 1440) %>% #assembly
  add_constraint(6*Chairs+4*Desks+Frames+25*Tables <= 2000) %>% #machining
  add_constraint(7*Chairs+10*Desks+2*Frames+12*Tables <= 2000) %>% #painting
  add_constraint(40*Chairs+25*Desks+5*Frames+16*Tables <= 9600) %>% #wood
 
  
  solve_model(with_ROI(solver = "glpk"))
  print(solver_status(result2))
  get_solution(result2, Chairs)
  get_solution(result2, Desks)
  get_solution(result2, Frames)
  get_solution(result2, Tables)
  print(objective_value(result2))
```

Well, it looks like it's not as simple as there being one bottleneck resource- if we have more painting hours, our model tells us to make a mix of frames and chairs, but still no tables or desks. I'm guessing that tables are too inefficient in machining and desks in assembly?